Different forms of uncertainty differently affect the evolution of social learning

### 2022-08-01 ms.tex
full revision with N=1000 data

\newcommand{\meanvar}[1]{\langle #1 \rangle}
\newcommand{\meansl}{\meanvar{s}}
\newcommand{\meanpi}{\meanvar{\pi}}
\newcommand{\meansoc}{\meanvar{\pi_\mathrm{S}}}
\newcommand{\meanasoc}{\meanvar{\pi_\mathrm{A}}}
\newcommand{\meanT}{\meanvar{T}}

@@

Social learning enhances problem solving when acquiring information from others is more efficient than learning on one's own~\cite{Laland2004}. However, social learning can also lead individuals astray if, for example, they are copying outdated information. Theoretical models have helped clarify the circumstances under which social learning should be evolutionarily favored ~\cite{BoydRicherson1985}. This theoretical work has motivated much empirical research and helped predict social learning behavior in humans and other species ~\cite{McElreath2005,Kendal2018,Allen2019}.

One proposal arising from this literature maintains that social learning is as an adaptation for dealing with uncertainty, and one that humans heavily rely on. Uncertainty weighs particularly heavily on human adaptation because of our long lifespans, cosmopolitan distribution, and the resulting necessity to adapt to relatively coarse-grain environments \cite{levins1962}. However, the term ``uncertainty" often conflates several different interpretations of that word. For example, different models define uncertainty in terms of the rate of environmental change, spatial environmental variation, or the reliability of information acquired from the environment. Furthermore, in translating mathematical models to verbal predictions, the formal meaning of terms like uncertainty is often lost \cite{lawson1988probability}, facilitating such conflation. As the number of formal models of social learning have expanded, an ever increasing number
of modeling choices ~\cite[Figure 1]{Kendal2018} and formalizations of uncertainty have made it difficult to compare across models and consolidate our understanding of the contexts in which social learning is adaptive. Finally, while risk and uncertainty are often used interchangeably to refer to non-deterministic outcomes, it can be important to distinguish a narrower definition of uncertainty wherein the probabilities of outcomes remain unknown \cite{knight1921, volz2012}.

Other (asocial) forms of learning may also be adaptations to uncertainty, but are usually under-emphasized in this literature. Because cultural evolutionary models are more often used to explain social learning mechanisms, they tend to use a single, and often under-powered individual-learning mechanism for comparison. This modelling decision may underestimate the extent to which individual learning can address the adaptive challenges solved by social learning, but also underestimate the quality of information available for social learning.

To address these concerns we developed an evolutionary agent-based model that simultaneously operationalizes several
different forms of uncertainty and endows agents with a powerful learning mechanism. We have chosen to model four kinds of uncertainty that are common in cultural evolution models and the related empirical literature:  (1) temporal environmental variability; (2) selection set size (the number of possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for different behaviors); and (4) the effective lifespan of agents (the number of opportunities for individual learning).\ps{citation needed}. The model also involves a more cognitively realistic and empirically-motivated learning mechanism, the softmax algorithm. This allows a more accurate assessment of tradeoffs between social and individual learning by fleshing out a learning mechanism that makes full use of the individually and socially learned data. We use this integrated model to ask what kinds of uncertainty are likely to favor social learning, and to clarify the logic of why this would be the case.

@@

Temporal environmental variability is fundamental to most evolutionary models of
social learning. Such fluctuations have been proposed as an important selective
pressure for learned as opposed to genetically fixed behaviors
~\cite{Richerson2000}. On the other hand, environments that change too quickly
will lead the population to have outdated, and therefore maladaptive information,
favoring individual updating over social learning point~\cite{Feldman1996,
BoydRicherson1985}. This suggests that intermediate levels of temporal variation
are important for the evolution of social learning mechanisms ~\cite{aoki2005}.

Despite the centrality of temporal fluctuations, even this construct is
implemented in several ways \cite{aoki2014evolution}. Most commonly, it is modeled
as an independent probability of the environment changing its state each
generation
~\cite{BoydRicherson1985,Rogers1988,Feldman1996,McElreath2005,Enquist2007,perreault2012bayesian,aoki2014evolution}.
The probabilities of environmental change are usually fixed per generation,
%~\cite{BoydRicherson1985,Rogers1988,kendal2009evolution}, but they have also been
modeled as deterministic cycles ~\cite{Feldman1996, aoki2014evolution}.
Furthermore, the consequences of environmental change can range from mild to
catastrophic. In the latter case, a change of environment results in the total
elimination of any adaptive behavior, which must be learned \emph{de novo} for the
new environment~\cite{Rogers1988}. Such catastrophic environmental changes present
a large adaptive challenge as individuals cannot bank on accumulated information
from previous generations. Other models of environmental change introduce less
uncertainty as they fluctuate between two environmental states with corresponding
adaptive behaviors. This means that all individuals with previously maladaptive
behaviors then have adaptive behaviors upon the environment changing
~\cite{perreault2012bayesian}. Which strategy one chooses for modelling temporal
change can have consequences for at least some of our inferences
~\cite{aoki2014evolution}.

\textbf{Selection set size.}
When one does not know which option to take, uncertainty increases with the number of options. This is a mathematical fact at the core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the selection set doubles in size  \cite{mackay2003information}. 
This is intuitive; the more options to consider, the greater one's uncertainty about which option is the best choice.

In studies of social learning, the size of the selection set is very often limited
to just two options.  For example, in empirical studies of \emph{bombus
terresteris} (bumble bee)~\cite{Baracchi2018} and \emph{parus major} (great
tit)~\cite{Aplin2017}, experimenters provided two behaviors for the bees or birds
to choose from, where one yielded a higher payoff than the other. %respectively,
each yielding greater or smaller payoffs depending on experimental treatment.
Similarly, many human studies have used just two or three possible
behaviors~\cite{McElreath2005,Toyokawa2019}. The simpler empirical choices are
often motivated by modeling studies with similar formulations ~\cite{boyd1995does,
Rogers1988,perreault2012bayesian}.\cm{pretty sure boyd1995 uses same trick as
rogers 88 so not sure why they were being used as illustrating separate kinds of
behavior selection sets. correct if wrong} Larger, or more open sets of behavioral
choices are not uncommon in experiments trying to capture more complex or
naturalistic tasks ~\cite{derex2013, wasielewski2014}, but these map on
imperfectly to the theoretical literature. Some models have studied systems with
larger, but defined, selection sets ~\cite{Rendell2010}, while in others the
number of options is subsumed under probabilities of getting a high payoff
~\cite{Enquist2007}. Researchers have also compared models with a different number
of environment states, corresponding to a different number of optimal behaviors
through time ~\cite{Feldman1996}, but it is very rare to find the size of the
selection set manipulated \emph{within} generations as a measure of changing
uncertainty in cultural evolutionary models.

@@

\textbf{Effective lifespan.} \mt{Restructure: lead with ``It stands to reason''
sentence or similar that gets to the point of this par more quickly, that lifespan
affects uncertainty. Also this Effective Lifespan section seems too long.}
Learning can be viewed as a process by which an individual reduces or otherwise
manages their uncertainty \cite{jacobs2011bayesian,clark2013whatever}. It is an
iterative process by which one repeatedly acquires information, forms
representations and predictions, and tests and refines those representations and
predictions. It therefore stands to reason that the more opportunities one has for
learning, the more one can in principle reduce one's uncertainty, assuming the
environment has not changed in the meantime. Similarly, a reduction in the number
of opportunities to learn will increase the uncertainty about whether one has
accurately learned about the available behavioral options and their associated
payoffs. We refer to the number of learning opportunities as an individual's
effective lifespan to imply that it is the number of opportunities to learn
throughout the individual's lifespan that matters here, not necessarily how long
they live in the absence of relevant learning opportunities.

Empirically, the number of learning opportunities can be manipulated in the lab,
but will also correspond with an individual's age the real world. In multi-round
studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds ~\cite{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain
about the task early on. In a more naturalistic context, Aplin et al. (2017)
\nocite{Aplin2017} found that younger great tits more readily used social
information compared with older individuals, possibly because they had accrued
less information via individual learning, and also possibly because younger
individuals have the most to gain (given their expected remaining lifespan) by
switching to superior behavioral options. 
% \ps{I got rid of the Leris and Reader reference because it didn't seem relevant
% to age-based uncertainty. They just showed that young guppies exposed to
% reliable models were more likely to use social information later. I also added
% some human studies, which seems important.} 
Cross-cultural studies on humans have shown the importance of childhood as a phase
of heavy social learning ~\cite{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \cite{kline2013teaching}.
%The age of individuals was found to have an effect on social learning in both
%\emph{parus major}~\cite{Aplin2017} and \emph{poecilia reticulata}
%(guppies)~\cite{Leris2016}. \ps{What effect? I'm not sure } \mt{Need better
%connection between lifespan and age, or better references to support use of
%lifespan}.
Individuals with shorter effective lifespans will end their lives with more
uncertainty about which behaviors are optimal, and thereby transmit some of that
uncertainty to anyone who learns from them. The effective lifespan for learning
varies across tasks, individuals, and species, yet most models assume one learning
opportunity per generation ~\cite{Feldman1996,Henrich1998, perreault2012bayesian}.
Some models do allow several cultural generations within 1 genetic generation
~\cite{Enquist2007}, but little formal theory explicitly examines the role of
uncertainty based on number of learning opportunities into our understanding of
the evolution of social learning.

@@

Many models of social learning assume little, or nothing, cognitive capabilities
of their agents. For example, a common modelling strategy compare the payoffs of
agents with different pure learning strategies (some will always engage in social
learning, and others in individual learning) while revealing nothing about the
cognition underlying individuals' decisions  ~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. Some more complicated learning strategies that make use of both
socially and individually learned information have been proposed
~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to
reflect cognitive mechanisms for integrating information. Though such
simplifications are often useful, several authors have warned of the risks to
\emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016,
Kendal2018}, given such details may be theoretically meaningful. In order to
understand how social learning evolves as an adaptation to uncertainty, we need to
endow agents with a  biologically plausible cognition capable of integrating
information from various sources in a payoff-maximizing way. \cm{I don't
  understand this next sentence, but think it can be cut with no loss of meaning:
  In our model we do not impose any structure in randomness between behaviors, so
our cognitive model is only sensitive to total randomness.} To do so, we let
agents see the payoffs to behaviors and use the softmax function to convert these
observations to a probability distribution from which they strategically select
their own behavior. \cm{does this still make sense? and does it need a citation?}

@@

We operationalized four separable varieties of uncertainty that affect the
ability of agents to find and choose the optimal behavior: 
\emph{environmental variability}; \emph{payoff ambiguity};
\emph{selection set size}; and the \emph{effective lifespan}. Environmental
variability is the probability that the optimal behavior changes
between generations. We do not consider within-generation variability. 
Furthermore, our model is not spatial, so agents have free, instant
access to all behaviors, and there is no spatial uncertainty. 
These conditions reflect humans in a computer-based social learning evolution
experiment~\cite{McElreath2005,Morgan2012,Derex2016}. Spatial dimensions could also be
considered irrelevant if costs associated with accessing a behavior were equal,
e.g., if it was assumed to be equally costly for animals to
access alternative food sources~\cite{Aplin2017,Baracchi2018}.  
% This is empirically justified since cultural evolution
% occurs when meta-cognitive strategies and information are transferred via
% learning and social influence, instead of by physical reproduction and death
% as in genetic evolution.  
The selection set size is the total number of possible
behaviors agents choose from, assumed to be constant within a simulation.  
Payoff ambiguity is the payoff difference between optimal
and non-optimal behaviors. The effective life span is the
number of time steps in which agents perform behaviors within a generation. 
Each simulated environment provides a distinct challenge
to the agents, which will lead agents to differently evolve to become social
learners or not, depending on which strategy optimizes net payoffs.

@@

To understand how different forms of uncertainty interact to affect the evolution of
social learning, we measured how frequently social learning evolved across
uncertainty parameters. We analyze outcomes by plotting the mean observed social
learning frequency, $\meansl$, across all agents and 1000 simulation trials for each
uncertainty parameter setting.   
% This is convenient because we are most certain about how $\meansl$ depends on $u$:
% we expect $\meansl$ to decrease as $u$ increases since environmental variability
% is well understood to cause socially-transmitted information to become outdated. 
To inspect patterns in $\meansl$ over all uncertainty parameters, we analyze nine
plots of $\meansl$ over $u$ in a $3\times3$ plot array
(Figure~\ref{fig:mainResults}). All plots show how $\meansl$ changes (y-axis)
with environmental variability, $u$ (x-axis). Selection set size, $B$ increases
from left to right in the plot array (columns). Payoff ambiguity via $\pilow$ 
increases from top to bottom in the array (recall $\pihigh=0.9$ for all analyses).
Effective lifespan, $L$, is indicated by plot line and marker colors. 
Several patterns in social learning prevalence emerged as 
the uncertainty parameters were varied (Figure~\ref{fig:mainResults}). 
% We support our conclusions regarding the evolution of social learning by analyzing
% the time to fixation and by comparing the average net payoffs in our simulations
% to the expected payoffs for individuals in both populations of all social learners
% and in populations of all individual learners. We thus use our analysis to
% understand changes to how $\meansl$ changes over $u$ for different uncertainty
% parameter settings.

First, we observed that greater $u$ uniformly decreased
$\meansl$ across all other uncertainty parameter values. The other uncertainty
parameters changed at which $u$ $\meansl$ began to decrease,
and how rapidly $\meansl$ decreased over $u$. For example, 
when payoff ambiguity and selection set size were
both small ($\pilow = 0.1$ and $B=2$), social learning evolved only when $u$ was
small. These parameter settings present a relatively easy problem for individuals.
Furthermore, it is relatively costly for agents to be misled by outdated social 
information in this case, especially for longer lifespans for finding the 
optimal behavior (Figure~\ref{fig:mainResults}, upper left). 
In another example, when $\pilow = 0.8$, $\meansl$ decreased less rapdily over $u$ due to 
weak selection caused by little advantage for performing the optimal versus
non-optimal behavior (see rows of Figure~\ref{fig:mainResults}).

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning, all other uncertainty parameters static. Environments with 
larger $B$ present a more difficult problem for trial-and-error learning within
a lifespan. This effect is amplified because a misidentified optimal
behavior is easier to correct for larger $B$ relative to smaller $B$.  
To understand this, consider the case where a social learner child learns
that behavior 1 has an expected payoff of 0.8, and the rest have an expected
payoff of 0, but the optimal behavior changes between generations. Assume
the child tries behavior 1, but it pays off 0.
If $B=2$, then at within-generation $t=2$, $\ledger = (0.4, 0.0)$. The probability
the agent explores behavior 2 is now 0.02. However, if $B=10$ the agent has
$\ledger = (0.4, 0.0, \ldots, 0.0)$; the probability
the agent explores one of the nine alternative behaviors increases to 0.14.
If the agent again tries behavior 1, the probability of exploring alternatives
with $B=2$ is 0.21, but with $B=10$ the probability of exploring alternatives
greatly increases to 0.71. This explanation does not apply when $L=1$, where
social learning is increasingly beneficial as $B$ increases because it is more 
valuable to be possibly misled by outdated information than it is for agents
to blindly guess which is optimal.

Payoff ambiguity had a modest effect when $\pilow = 0.1$ increased to $\pilow =
0.45$, causing social learning to be favored across more values of $u$, and across
all selection set sizes and lifespans (first two rows of
Figure~\ref{fig:mainResults}).  This is because increased payoff ambiguity increases
the difficulty, and reduces the overall benefit, of discerning which behavior is
optimal; it is worth the risk for agents to possibly learn outdated information.
When $\pilow = 0.8$ the largest effect is the smoothing and flattening of the
$\meansl$ curves, which indicates more drift (bottom row of
Figure~\ref{fig:mainResults}).  For smaller values of $u$, this results in less
frequent evolution of social learning for the same values of $u$ and $B$. But after
the value of $u$ at which $\meansl \approx 0.5$, social learning evolves more
frequently than for the same $u$, $B$, and $L$. When $\pilow \to \pihigh$, agents 
accumulate increasingly more payoffs, whether they are social learners or not. 
Greater payoff ambiguity means lower risk to try possibly outdated social
information, search for the true optimal if the learned one fails.

When $L$ is smaller, agents have fewer opportunities to learn through trial and
error, so it is more difficult for them to discern which strategy is optimal, social
or asocial. We observed two distinct trends in the evolution of social learning as
$L$ was varied.  First, when $\pilow$ was sufficiently small ($\pilow=0.1, 0.45$),
we observed that shorter lifespan selected for social learning for some values of
$u$.  When $\pilow=0.8$, shorter lifespans decreased $\meansl$ in some cases and
increased $\meansl$ in other cases due to greater evolutionary drift.  In the first
case, $\meansl$ was reduced for $L=1$ for small values of $u$, when other $L > 1$
led to $\meansl \approx 1$. At larger $u$ , the $L=1$ case yields non-zero
$\meansl$, while for the other $L > 1$, $\meansl \approx 0$.

@@

\subsection{Supplemental measures of social learning}

\mt{This subsec needs revision and associated figures need to be updated with 
$N=1000$ data}

>>>>

\newcommand{\meanvar}[1]{\langle #1 \rangle}
\newcommand{\meansl}{\meanvar{s}}
\newcommand{\meanpi}{\meanvar{\pi}}
\newcommand{\meansoc}{\meanvar{\pi_\mathrm{S}}}
\newcommand{\meanasoc}{\meanvar{\pi_\mathrm{A}}}
\newcommand{\meanT}{\meanvar{T}}
\newcommand{\meanG}{\meanvar{G}}

@@

Social learning enhances problem solving when acquiring information from others is more
efficient than learning on one's own~\cite{Laland2004}. However, social learning can also
lead individuals astray if, for example, they are copying outdated information.
Theoretical models have helped clarify the circumstances under which social learning
should be evolutionarily favored ~\cite{BoydRicherson1985}. This theoretical work has
motivated much empirical research and helped predict social learning behavior in humans
and other species ~\cite{McElreath2005,Kendal2018,Allen2019}.  This literature
demonstrates that social learning is a critical adaptation found across taxa for dealing
with uncertainty.  Uncertainty weighs particularly heavily on human adaptation because of
our long lifespans, cosmopolitan distribution, and the resulting necessity to adapt to
relatively coarse-grain environments \cite{levins1962}. However, the term ``uncertainty"
often conflates several different interpretations of that word. For example, different
models define uncertainty in terms of the rate of environmental change, spatial
environmental variation, or the reliability of information acquired from the environment.
Furthermore, in translating mathematical models to verbal predictions, the formal meaning
of terms like uncertainty is often lost \cite{lawson1988probability}, facilitating such
conflation. Furthermore, it can be important to distinguish a narrower definition of
uncertainty wherein the probabilities of outcomes remain unknown~\cite{knight1921,
volz2012}. As the number of formal models of social learning have expanded, an ever
increasing number of modeling choices ~\cite[Figure 1]{Kendal2018} and formalizations of
uncertainty have made it difficult to compare across models and consolidate our
understanding of the contexts in which social learning is adaptive. 
% Finally, while risk and uncertainty are often used interchangeably to refer to
% non-deterministic outcomes,

Other (asocial) forms of learning may also be adaptations to uncertainty that could be
accounted for, but are typically glossed over.  In general, learning is an iterative
process by which one repeatedly acquires information, forms representations and
predictions, and tests and refines those representations and predictions to manage
uncertainty~\cite{jacobs2011bayesian,clark2013whatever}.  Because cultural evolutionary
models are more often used to explain social learning mechanisms, they tend to use a
single, weak individual-learning mechanism for comparison. Abstracting away asocial forms
of learning may lead us to underestimate the extent to which individual learning provides
quality information for social learning and for the individual to adapt to the challenges
assumed to be solved by social learning. 
% underestimate the quality of information
% available for social learning.

To address these concerns we developed an evolutionary agent-based model that
simultaneously operationalizes several different forms of uncertainty and endows agents
with a relatively powerful learning mechanism. We have chosen to model four kinds of
uncertainty that are common in cultural evolution models and the related empirical
literature:  (1) temporal environmental variability; (2) selection set size (the number of
possible behaviors); (3) payoff ambiguity (the difference in the expected rewards for
different behaviors); and (4) the effective lifespan of agents (the number of
opportunities for individual learning), explained below. The model also assumes a more
cognitively realistic and empirically-motivated learning mechanism, the softmax algorithm.
This allows for a more accurate assessment of tradeoffs between 
social and asocial learning
by fleshing out an individual-level learning mechanism that makes full use of 
the individually and socially learned data. 
In this work, social learners learn from the previous generation, and 
asocial learners do not. All agents engage in individual-level learning within
their lifespans whether they are social learners or not.
We use this integrated model to ask what kinds of uncertainty are likely to
favor social learning, and to clarify the logic of why this would be the case.

@@

Temporal environmental variability is fundamental to most evolutionary models of
social learning. Such fluctuations have been proposed as an important selective
pressure for learned as opposed to genetically fixed behaviors
~\cite{Richerson2000}. On the other hand, environments that change too quickly
will lead the population away from social learning to avoid outdated, 
and therefore maladaptive information~\cite{Feldman1996,
BoydRicherson1985}. This suggests that intermediate levels of temporal variation
are important for the evolution of social learning mechanisms~\cite{aoki2005}.

Despite the theoretical centrality of temporal environmental variability, this construct is
implemented in several ways~\cite{aoki2014evolution}. Most commonly, it is modeled
as an independent probability of the environment changing its state each
generation
~\cite{BoydRicherson1985,Rogers1988,Feldman1996,McElreath2005,Enquist2007,perreault2012bayesian,aoki2014evolution}.
% The probabilities of environmental change are usually fixed per generation
% ~\cite{BoydRicherson1985,Rogers1988,kendal2009evolution}, 
but has also been modeled as deterministic cycles~\cite{Feldman1996, aoki2014evolution}.
Furthermore, the consequences of environmental change can range from mild to
catastrophic. In the latter case, a change of environment results in the total
elimination of any adaptive behavior, which must be learned \emph{de novo} for the
new environment~\cite{Rogers1988}. Such catastrophic environmental changes present
a large adaptive challenge as individuals cannot rely on accumulated information
from previous generations. Other models of environmental change introduce less
uncertainty as they fluctuate between two environmental states with corresponding
adaptive behaviors. This means that previously maladaptive
behaviors become adaptive when the environment changes
~\cite{perreault2012bayesian}. The chosen mechanism for modelling temporal
change has real theoretical consequences~\cite{aoki2014evolution}.

\textbf{Selection set size.}
When one does not know which option to take, uncertainty increases with the number of options, which we call the selection set size.
% This is a mathematical fact at the
% core of information theory: the number of bits required to reliably encode the optimal solution increases by one every time the
% selection set doubles in size~\cite{mackay2003information}. 
% This is intuitive; the more options to consider, the greater one's uncertainty about which option is the best choice. 
In studies of social learning, the selection set is often limited
to just two options.  For example, in empirical studies of \emph{bombus
terresteris} (bumble bee)~\cite{Baracchi2018} and \emph{parus major} (great
tit)~\cite{Aplin2017}, experimenters provided two behaviors for the bees or birds
to choose from, where one yielded a higher payoff than the other. %respectively,
each yielding greater or smaller payoffs depending on experimental treatment.
Similarly, many human studies have used just two or three possible
behaviors~\cite{McElreath2005,Toyokawa2019}. Behavioral study designs are
often motivated by modeling studies with similar formulations~\cite{Rogers1988,boyd1995does,
perreault2012bayesian}.
% \cm{pretty sure boyd1995 uses same trick as
% rogers 88 so not sure why they were being used as illustrating separate kinds of
% behavior selection sets. correct if wrong} 
Larger, more open sets of behavioral
choices are not uncommon in experiments trying to capture more complex or
naturalistic tasks~\cite{derex2013, wasielewski2014}, but these map on
imperfectly to the theoretical literature \mt{How so?}. Some models have studied systems with
larger, but defined, selection sets~\cite{Rendell2010}, while in others the
number of options is implicit in the model payoff structure
~\cite{Enquist2007} \mt{How does this work exactly in Enquist et al 2007?}. 
Researchers have also compared models with a different number
of environment states, corresponding to a different number of optimal behaviors
through time ~\cite{Feldman1996}. These examples share one common feature: 
the size of the selection set was held static \emph{within} model generations.
% as a measure of changing uncertainty in cultural evolutionary models.

@@

\textbf{Effective lifespan.} 
The more opportunities one has for
learning in their lifespan, the more uncertainty can be reduced, assuming a
static environment. Similarly, a reduction in the number
of opportunities to learn will increase the uncertainty about whether one has
accurately learned about the available behavioral options and their associated
payoffs. That uncertainty is passed on to future generations.
% Learning is an iterative process by which one repeatedly acquires information, forms
% representations and predictions, and tests and refines those representations and
% predictions to manage uncertainty \cite{jacobs2011bayesian,clark2013whatever}. 
We refer to the number of learning opportunities as an individual's
effective lifespan to imply that it is the number of opportunities to learn
throughout the individual's lifespan that matters here, not necessarily how long
they live in the absence of relevant learning opportunities.

Empirically, the number of learning opportunities can be manipulated in the lab,
but will also correspond with an individual's age the real world. In multi-round
studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds ~\cite{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain
about the task early on. In a more naturalistic context, Aplin et al. (2017)
\nocite{Aplin2017} found that younger great tits more readily used social
information compared with older individuals, possibly because they had accrued
less information via individual learning, and also possibly because younger
individuals have the most to gain (given their expected remaining lifespan) by
switching to superior behavioral options. 
% \ps{I got rid of the Leris and Reader reference because it didn't seem relevant
% to age-based uncertainty. They just showed that young guppies exposed to
% reliable models were more likely to use social information later. I also added
% some human studies, which seems important.} 
Cross-cultural studies on humans have shown the importance of childhood as a phase
of heavy social learning ~\cite{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \cite{kline2013teaching}.
%The age of individuals was found to have an effect on social learning in both
%\emph{parus major}~\cite{Aplin2017} and \emph{poecilia reticulata}
%(guppies)~\cite{Leris2016}. \ps{What effect? I'm not sure } \mt{Need better
%connection between lifespan and age, or better references to support use of
%lifespan}.
% Individuals with shorter effective lifespans will end their lives with more
% uncertainty about which behaviors are optimal, and thereby transmit some of that
% uncertainty to anyone who learns from them. 
The effective lifespan for learning
varies across tasks, individuals, and species, yet most models assume one learning
opportunity per generation ~\cite{Feldman1996,Henrich1998, perreault2012bayesian}.
Some models do allow several cultural generations within 1 genetic generation
~\cite{Enquist2007}, but little formal theory explicitly examines the role of
uncertainty based on number of learning opportunities into our understanding of
the evolution of social learning.

@@

Many models of social learning assume little, or nothing, cognitive capabilities
of their agents. For example, a common modelling strategy compare the payoffs of
agents with different pure learning strategies (some will always engage in social
learning, and others in individual learning) while revealing nothing about the
cognition underlying individuals' decisions  ~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. Some more complicated learning strategies that make use of both
socially and individually learned information have been proposed
~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to
reflect cognitive mechanisms for integrating information. Though such
simplifications are often useful, several authors have warned of the risks to
\emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016,
Kendal2018}, given such details may be theoretically meaningful. In order to
understand how social learning evolves as an adaptation to uncertainty, we need to
endow agents with a biologically plausible cognition capable of integrating
information from various sources in a payoff-maximizing way~\cite{Gershman2019}. 
% \cm{I don't
%   understand this next sentence, but think it can be cut with no loss of meaning:
%   In our model we do not impose any structure in randomness between behaviors, so
% our cognitive model is only sensitive to total randomness.} 
To do so, we let
agents see the payoffs to behaviors and use the softmax function to convert these
observations to a probability distribution from which they strategically select
their own behavior. %\cm{does this still make sense? and does it need a citation?}

@@

We operationalized four separable varieties of uncertainty that affect the
ability of agents to find and choose the optimal behavior: 
\emph{environmental variability}; 
\emph{selection set size}; \emph{payoff ambiguity}; and the \emph{effective lifespan}. 
Environmental variability is the probability that the optimal behavior changes
between generations. We do not consider within-generation variability. 
Furthermore, our model is not spatial, so agents have free, instant
access to all behaviors, and there is no spatial uncertainty. 
These conditions reflect humans in a computer-based social learning evolution
experiment~\cite{McElreath2005,Morgan2012,Derex2016}. Spatial dimensions could also be
considered irrelevant if costs associated with accessing a behavior were equal,
e.g., if it was assumed to be equally costly for animals to
visit alternative food sources~\cite{Aplin2017,Baracchi2018}.  
% This is empirically justified since cultural evolution
% occurs when meta-cognitive strategies and information are transferred via
% learning and social influence, instead of by physical reproduction and death
% as in genetic evolution.  
The selection set size is the total number of possible
behaviors agents choose from, assumed to be constant within a simulation.  
Payoff ambiguity is the payoff difference between optimal
and non-optimal behaviors. The effective life span is the
number of time steps in which agents perform behaviors within a generation. 
Each simulated environment provides a distinct challenge
to the agents, which will lead agents to differently evolve to become social
learners or not, depending on which strategy optimizes net payoffs.

@@

To understand how different forms of uncertainty interact to affect the evolution of
social learning, we measured how frequently social learning evolved across
uncertainty parameters. We analyzed outcomes by plotting the mean observed social
learning frequency, $\meansl$, with mean taken across all agents and 1000 simulation trials for each
uncertainty parameter setting.   
% This is convenient because we are most certain about how $\meansl$ depends on $u$:
% we expect $\meansl$ to decrease as $u$ increases since environmental variability
% is well understood to cause socially-transmitted information to become outdated. 
% To inspect patterns in $\meansl$ over all uncertainty parameters, 
We analyze nine plots of how $\meansl$ (y-axis in each plot) changes over $u$ (x-axis) 
in a $3\times3$ plot array (Figure~\ref{fig:mainResults}). 
% All plots show how $\meansl$ changes (y-axis)
% with environmental variability, $u$ (x-axis). 
Selection set size, $B$ increases from left to right in the plot array (columns). 
Payoff ambiguity increases as $\pilow$ increases from top to bottom in the array 
(recall $\pihigh=0.9$ for all analyses).
Effective lifespan, $L$, is indicated by line and marker colors (inset). 
Several patterns in social learning prevalence emerged as 
these uncertainty parameters were varied (Figure~\ref{fig:mainResults}). 
% We support our conclusions regarding the evolution of social learning by analyzing
% the time to fixation and by comparing the average net payoffs in our simulations
% to the expected payoffs for individuals in both populations of all social learners
% and in populations of all individual learners. We thus use our analysis to
% understand changes to how $\meansl$ changes over $u$ for different uncertainty
% parameter settings.

First, we confirmed the current theoretical understanding that greater $u$ uniformly decreased
$\meansl$ across all other uncertainty parameter values. The other uncertainty
parameters changed the $u$ value at which $\meansl$ began to decrease,
and how rapidly $\meansl$ decreased over $u$. 
For example, when payoff ambiguity and selection set size were
both small ($\pilow = 0.1$ and $B=2$), social learning evolved only when $u$ was
small. 
% These parameter settings present a relatively easy problem for individuals.
% Furthermore, it is relatively costly for agents to be misled by outdated social 
% information in this case, especially for longer lifespans for finding the 
% optimal behavior (Figure~\ref{fig:mainResults}, upper left). 
When $\pilow = 0.8$ and $B=2$, for exapmle, $\meansl$ decreased less rapdily over $u$ due to 
drift caused by weak selection caused by little advantage for performing the optimal versus
non-optimal behavior (see rows of Figure~\ref{fig:mainResults}). We analyzed several
such patterns of social learning evolution and suppression across different uncertainty settings.

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning, with all other uncertainty parameters held constant. This is first 
because it is more difficuilt to find the optimal behavior with a larger $B$.
% more options present a more difficult problem for trial-and-error learning within
% a lifespan. 
This effect is amplified because agents in a world with larger $B$ are more
cognitively flexible compared to those in a smaller $B$ world. 
When $B$ is larger, agents are relatively more likely to choose an alternative to
their current known optimum. %is easier to correct for larger $B$ relative to smaller $B$.  
To understand this, consider the case where a social learner child learns
that behavior 1 has an expected payoff of 0.8, and the rest have an expected
payoff of 0, but the optimal behavior changed between generations. Assume
the child tries behavior 1, but it pays off 0.
If $B=2$, then at the next within-generation time step, $\ledger = (0.4, 0.0)$. The probability
the agent explores behavior 2 is now 0.02. However, if $B=10$ the agent has
$\ledger = (0.4, 0.0, \ldots, 0.0)$; the probability
the agent explores one of the nine alternative behaviors increases to 0.14.
If the agent again tries behavior 1, the probability of exploring alternatives
with $B=2$ is 0.21, but with $B=10$ the probability of exploring alternatives
greatly increases to 0.71. This explanation does not apply when $L=1$, where
individual learning does not occur. %social learning is the only possible form of learning.

Payoff ambiguity had a modest effect when $\pilow = 0.1$ increased to $\pilow =
0.45$, causing social learning to be favored across more values of $u$ (first two rows of
Figure~\ref{fig:mainResults}).  This is because increased payoff ambiguity increases
the difficulty, and reduces the overall benefit, of discerning which behavior is
optimal.  When $\pilow = 0.8$ the
$\meansl$ curves become smoother, indicating more evolutionary drift towards 
possibly non-optimal fixations (bottom row of
Figure~\ref{fig:mainResults}). For smaller values of $u$, this results in less
frequent evolution of social learning for the same values of $u$ and $B$. But past
the value of $u$ where $\meansl \approx 0.5$ this switches, and social learning evolves more
frequently than for smaller $\pilow$. 
% When $\pilow \to \pihigh$, agents 
% accumulate increasingly more payoffs, whether they are social learners or not. 
Greater payoff ambiguity means lower risk to try possibly outdated social
information by reducing the penalty for performing non-optimal behaviors.

We observed two distinct trends in the evolution of social learning as
$L$ was varied.  First, when $\pilow$ was sufficiently small ($\pilow=0.1, 0.45$),
we observed that shorter lifespan extended the evolution of social learning over more values of
$u$, since payoff-biased social learning means teachers are reliable.
When $\pilow=0.8$, shorter lifespans decreased $\meansl$ in some cases and
increased $\meansl$ in other cases due to greater evolutionary drift. In
this setting, many agents will obtain a payoff of 1 at each time step in their 
lifetime, which, combined with finite population size, will make it more likely for 
evolution to settle on a non-optimal population fixation.
% In the first
% case, $\meansl$ was reduced for $L=1$ for small values of $u$, when other $L > 1$
% led to $\meansl \approx 1$. At larger $u$ , the $L=1$ case yields non-zero
% $\meansl$, while for the other $L > 1$, $\meansl \approx 0$. 
Smaller $L$ means agents end their lives more uncertain about which behavior was
optimal, but payoff-biased teacher selection and learning can remedy this if
payoff ambiguity is sufficiently small.
% When $L$ is smaller, agents have fewer opportunities to learn through trial and
% error, so it is more difficult for them to discern which strategy is optimal, social
% or asocial.

@@

\subsection{The relative benefit of social learning and the presence of evolutionary drift}
